\textbf{Species and genus plant classification with NEON hyperspectral remote sensing data}

### 2020-11-05 R/classify_species.Rmd
combined var imp plots into one figure
MDA for both species and genus

library(tidyverse)
library(tidyr)
library(dplyr)
library(randomForest)
library(caret)
library(rfPermute)
library(parallel)
library(doParallel)

@@

rf_valid_results %>% 
  # count number of each obs/pred combination
  count(obs,pred) %>% 
  # fill in combinations with counts of zero
  complete(obs, pred, fill = list(n = 0)) %>% 
  # add rows witth species combinations not in predicted set
  right_join(all_combos) %>% 
  # set NA values to 0 in the n column
  mutate(n = ifelse(is.na(n), 0, n)) %>% 
  # set raster cell fill color based on count
  ggplot(aes(x = obs, y = fct_rev(pred), fill = n)) + 
  geom_raster() + 
  # add boxes around diagonal cells
  geom_tile(aes(color = obs == pred, width = 0.95, height = 0.95), size = 0.5) + 
  scale_color_manual(values = c(NA, "black")) + 
  # rotate x axis text 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "Truth", 
       y = "Predicted"
       #, title = "RandomForest Classification Confusion Matrix: Predicted vs. True SPECIES for Validation Set"
       ) + 
  scale_fill_gradient(low = "white", high = "#43a2ca") + 
  # add count in each cell
  geom_text(aes(label = n, alpha = n > 0)) + 
  # remove legend entirely
  theme(legend.position = "none",
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 16))

@@

```{r varImp_byWavelength_plot, fig.height=6, fig.width=10}
varImp_MDA_species <- ggplot(imp, aes(x=wavelength_nm,  
                y=MeanDecreaseAccuracy)) + 
  # add bars where each height is equal to the variable importance metric
  geom_col(aes(fill = MeanDecreaseAccuracy), color = NA) +
  theme_minimal() + 
  labs(x = "Wavelength (nm)\n", 
       y = "Variable importance \nMean Decrease in Accuracy (MDA)\n",
       #title = "Variable Importance: Mean Decrease Accuracy (MDA)",
       fill = "MDA")  + 
  scale_fill_gradient(low = "#d9d9d9", high = "#43a2ca") + 
   theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16))

@@

# save confusion matrix to file 
ggsave(filename = here::here("analysis","figures", "varImp_species.pdf"), width = 10, height = 10)
```

@@

rf_valid_results_genus %>% 
  # count number of each obs/pred combination
  count(obs,pred) %>% 
  # fill in combinations with counts of zero
  complete(obs, pred, fill = list(n = 0)) %>% 
  # add rows witth species combinations not in predicted set
  right_join(all_combos_genus) %>% 
  # set NA values to 0 in the n column
  mutate(n = ifelse(is.na(n), 0, n)) %>% 
  # set raster cell fill color based on count
  ggplot(aes(x = obs, y = fct_rev(pred), fill = n)) + 
  geom_raster() + 
  # add boxes around diagonal cells
  geom_tile(aes(color = obs == pred, width = 0.95, height = 0.95), size = 0.5) + 
  scale_color_manual(values = c(NA, "black")) + 
  # rotate x axis text 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "Truth", y = "Predicted", title = "RandomForest Classification Confusion Matrix: Predicted vs. True GENUS for Validation Set") + 
  scale_fill_gradient(low = "white", high = "#fc8d59") + 
  # add count in each cell
  geom_text(aes(label = n, alpha = n > 0)) + 
  # remove the color legend
  #guides(color = "none")
  # remove legend entirely
  theme(legend.position = "none")

@@

```{r varImp_MDA_genus, fig.height=6, fig.width=10}
ggplot(imp_genus, aes(x=wavelength_nm,  
                y=MeanDecreaseAccuracy)) + 
  # add bars where each height is equal to the variable importance metric
  geom_col(aes(fill = MeanDecreaseAccuracy), color = NA) +
  theme_minimal() + 
  labs(x = "Wavelength (nm)", 
       y = "Variable importance (MDA)",
       title = "GENUS classification variable importance: Mean Decrease Accuracy (MDA) vs. Wavelength",
       fill = "MDA")  + 
  scale_fill_gradient(low = "#d9d9d9", high = "#fc8d59")
```

```{r varImp_Gini_genus, fig.height=6, fig.width=10}
ggplot(imp_genus, aes(x = wavelength_nm,  
                y = MeanDecreaseGini)) + 
  # add bars where each height is equal to the variable importance metric
  geom_col(aes(fill = MeanDecreaseGini), color = NA) +
  theme_minimal() + 
  labs(x = "Wavelength (nm)", 
       y = "Variable importance (MDG)",
       title = "GENUS classification variable importance: Mean Decrease Gini (MDG) vs. Wavelength",
       fill = "MDG")  + 
  scale_fill_gradient(low = "#d9d9d9", high = "#fc8d59")
```

@@

```{r varImp_with_refl_spectrum}
imp_and_refl <- imp

```{r randomForest_varImpPlot_ggplot, eval=FALSE, include=FALSE}

# default varImpPlot visualization in randomForest
#varImpPlot(rf_classifier)

# format the randomForest variable importance for ggplot
imp <- as.data.frame(varImpPlot(rf_classifier))
imp$varnames <- rownames(imp) # row names to column
rownames(imp) <- NULL

# keep the n features with the highest MeanDecreaseAccuracy values 
imp_top30_mda <- arrange(imp, desc(MeanDecreaseAccuracy)) %>% 
  select(c(MeanDecreaseAccuracy, varnames)) %>% 
  slice_head(n = 40)

ggplot(imp_top30_mda, aes(x=reorder(varnames, MeanDecreaseAccuracy), 
                y=MeanDecreaseAccuracy
                #,color=MeanDecreaseAccuracy
                )) + 
  geom_point() +
  geom_segment(aes(x=varnames,xend=varnames,y=0,yend=MeanDecreaseAccuracy)) +
  #scale_color_continuous(type = "viridis") +
  ylab("MDA") +
  xlab("Variable Name") +
  ggtitle("RandomForest variable importance: Mean Decrease Accuracy (MDA)") +
  labs(color = "MDA") + 
  coord_flip() + 
  theme_minimal()
```

>>>>

library(tidyverse)
library(tidyr)
library(dplyr)
library(randomForest)
library(caret)
library(rfPermute)
library(parallel)
library(doParallel)
library(ggpubr)

@@

rf_valid_results %>% 
  # count number of each obs/pred combination
  count(obs,pred) %>% 
  # fill in combinations with counts of zero
  complete(obs, pred, fill = list(n = 0)) %>% 
  # add rows witth species combinations not in predicted set
  right_join(all_combos) %>% 
  # set NA values to 0 in the n column
  mutate(n = ifelse(is.na(n), 0, n)) %>% 
  # set raster cell fill color based on count
  ggplot(aes(x = obs, y = fct_rev(pred), fill = n)) + 
  geom_raster() + 
  # add boxes around diagonal cells
  geom_tile(aes(color = obs == pred, width = 0.95, height = 0.95), size = 0.5) + 
  scale_color_manual(values = c(NA, "black")) + 
  # rotate x axis text 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "True Species", 
       y = "Predicted Species"
       #, title = "RandomForest Classification Confusion Matrix: Predicted vs. True SPECIES for Validation Set"
       ) + 
  scale_fill_gradient(low = "white", high = "#43a2ca") + 
  # add count in each cell
  geom_text(aes(label = n, alpha = n > 0)) + 
  # remove legend entirely
  theme(legend.position = "none",
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 16))

@@

```{r varImp_byWavelength_plot, fig.height=6, fig.width=10}
varImp_MDA_species <- ggplot(imp, aes(x=wavelength_nm,  
                y=MeanDecreaseAccuracy)) + 
  # add bars where each height is equal to the variable importance metric
  geom_col(aes(fill = MeanDecreaseAccuracy), color = NA) +
  theme_minimal() + 
  labs(x = "Wavelength (nm)\n", 
       y = "Variable importance \n",
       title = "Variable Importance: Mean Decrease Accuracy (MDA) \n\nSpecies classification",
       fill = "MDA")  + 
  scale_fill_gradient(low = "#d9d9d9", high = "#43a2ca") + 
   theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        title = element_text(size = 16))

@@

# save confusion matrix to file 
#ggsave(filename = here::here("analysis","figures", "varImp_species.pdf"), width = 10, height = 10)
```

@@

rf_valid_results_genus %>% 
  # count number of each obs/pred combination
  count(obs,pred) %>% 
  # fill in combinations with counts of zero
  complete(obs, pred, fill = list(n = 0)) %>% 
  # add rows witth species combinations not in predicted set
  right_join(all_combos_genus) %>% 
  # set NA values to 0 in the n column
  mutate(n = ifelse(is.na(n), 0, n)) %>% 
  # set raster cell fill color based on count
  ggplot(aes(x = obs, y = fct_rev(pred), fill = n)) + 
  geom_raster() + 
  # add boxes around diagonal cells
  geom_tile(aes(color = obs == pred, width = 0.95, height = 0.95), size = 0.5) + 
  scale_color_manual(values = c(NA, "black")) + 
  # rotate x axis text 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "True Genus", 
       #title = "RandomForest Classification Confusion Matrix: Predicted vs. True GENUS for Validation Set",
       y = "Predicted Genus") + 
  scale_fill_gradient(low = "white", high = "#fc8d59") + 
  # add count in each cell
  geom_text(aes(label = n, alpha = n > 0)) + 
  # remove the color legend
  #guides(color = "none")
  # remove legend entirely
  theme(legend.position = "none",
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 14))

@@

```{r varImp_MDA_genus, fig.height=6, fig.width=10}
varImp_MDA_genus <- ggplot(imp_genus, aes(x=wavelength_nm,  
                y=MeanDecreaseAccuracy)) + 
  # add bars where each height is equal to the variable importance metric
  geom_col(aes(fill = MeanDecreaseAccuracy), color = NA) +
  theme_minimal() + 
  labs(x = "Wavelength (nm)\n", 
       y = "Variable importance\n",
       title = "Genus classification",
       fill = "MDA")  + 
  scale_fill_gradient(low = "#d9d9d9", high = "#fc8d59") + 
   theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        title = element_text(size = 16))

varImp_MDA_genus

```{r varImp_Gini_genus, fig.height=6, fig.width=10}
ggplot(imp_genus, aes(x = wavelength_nm,  
                y = MeanDecreaseGini)) + 
  # add bars where each height is equal to the variable importance metric
  geom_col(aes(fill = MeanDecreaseGini), color = NA) +
  theme_minimal() + 
  labs(x = "Wavelength (nm)", 
       y = "Variable importance (MDG)",
       title = "GENUS classification variable importance: Mean Decrease Gini (MDG) vs. Wavelength",
       fill = "MDG")  + 
  scale_fill_gradient(low = "#d9d9d9", high = "#fc8d59") + 
   theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16))
```

Single figure with MDA variable importance plots for both species and genus classifiers:

```{r varmImp_species}
ggpubr::ggarrange(varImp_MDA_species, varImp_MDA_genus, nrow = 2)

# save confusion matrix to file 
ggsave(filename = here::here("analysis","figures", "varImp.pdf"), width = 10, height = 10)
```

@@

### 2020-11-05 analysis/paper/paper.Rmd
combined var imp plots into one figure
MDA for both species and genus

Variable importance for species classification (Figure \@ref(fig:fig-varImpSpecies))

@@

```{r fig-varImpSpecies, fig.cap ="Variable importance for species classification", echo=FALSE,  out.width = "450px", message = FALSE, warning=FALSE}
library(here)
knitr::include_graphics(here::here("analysis","figures", "varImp_species.pdf"))
```

>>>>

Variable importance for species and genus classifiers (Figure \@ref(fig:fig-varImp))

@@

```{r fig-varImp, fig.cap ="Variable importance", echo=FALSE,  out.width = "450px", message = FALSE, warning=FALSE}
library(here)
knitr::include_graphics(here::here("analysis","figures", "varImp.pdf"))
```
